Appendix A

### 2023-11-10 manuscript/manuscript_appendix_a.Rmd
Update manuscript_appendix_a.Rmd

---
title: "Appendix A"
author: "Alex Zwanenburg"
date: "`r Sys.Date()`"
output: html_document
---

@@

```{=latex}
\begin{equation}
\phi_{\text{BC}}^{\lambda, x_0, s} (x_i) = 
\begin{cases}
\left( \left(\frac{x_i - x_0}{s} \right)^\lambda - 1 \right) / \lambda & \text{if } \lambda \neq 0\\
\log(\frac{x_i - x_0}{s}) & \text{if } \lambda = 0
\end{cases}
\end{equation}
```
Where $x_i - x_0 > 0$.
The location and scale-invariant Yeo-Johnson transformation is:

```{=latex}
\begin{equation}
\phi_{\text{YJ}}^{\lambda, x_0, s} (x_i) = 
\begin{cases}
\left( \left( 1 + \frac{x_i - x_0}{s}\right)^\lambda - 1\right) / \lambda & \text{if } \lambda \neq 0 \text{ and } x_i - x_0 \geq 0\\
\log(1 + \frac{x_i - x_0}{s}) & \text{if } \lambda = 0 \text{ and } x_i - x_0 \geq 0\\
-\left( \left( 1 - \frac{x_i - x_0}{s}\right)^{2 - \lambda} - 1 \right) / \left(2 - \lambda \right) & \text{if } \lambda \neq 2 \text{ and } x_i - x_0 < 0\\
-\log(1 - \frac{x_i - x_0}{s}) & \text{if } \lambda = 2 \text{ and } x_i - x_0 < 0
\end{cases}
\end{equation}
```

@@

Since in our situation $\{x_1, \ldots, x_n\}$ in $f^{\lambda, x_0, s}(x_1, \ldots, x_n)$ are considered fixed (i.e., known), $f^{\lambda, x_0, s}(x_1, \ldots, x_n)$ may be considered a likelihood function.
The log-likelihood function $\mathcal{l}^{\lambda, x_0, s}$ is then:
```{=latex}
\begin{equation}
\begin{split}
\mathcal{l}^{\lambda, x_0, s} & = \log f^{\lambda, x_0, s}(x_1, \ldots, x_n) \\
 & = \log f^{\lambda, x_0, s} (\phi^{\lambda, x_0, s}(x_1), \ldots, \phi^{\lambda, x_0, s}(x_n)) + \log \left|\mathbf{J}\right| \\
 & = \log f^{\lambda, x_0, s} (\phi^{\lambda, x_0, s}(x_1), \ldots, \phi^{\lambda, x_0, s}(x_n)) + \log \prod_{i=1}^n \frac{\partial}{\partial x_i} \phi^{\lambda, x_0, s}(x_i) \\
 & = -\frac{n}{2} \log{2 \pi \sigma^2} -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi^{\lambda, x_0, s}(x_i) - \mu \right)^2 + \sum_{i=1}^n \log \frac{\partial}{\partial x_i} \phi^{\lambda, x_0, s}(x_i)
\end{split}
\end{equation}
```
With $\mu$ the average of $\phi^{\lambda, x_0, s}(\mathbf{X})$ and $\sigma^2$ its variance.
The first two terms derive directly from the log-likelihood function of a normal distribution, and are not specific to the type of power transformation used.
However, the final term differs between Box-Cox and Yeo-Johnson transformations.

@@

Thus the final term is:
```{=latex}
\begin{equation}
\begin{split}
\sum_{i=1}^n \log \frac{\partial}{\partial x_i} \phi_{\text{BC}}^{\lambda, x_0, s}(x_i) & = 
\end{split}
\end{equation}
```

@@

>>>>

---
title: "Appendix A"
author: "Alex Zwanenburg"
date: "`r Sys.Date()`"
output: html_document
header-includes:
- \usepackage{amsmath}
- \usepackage{booktabs}
- \DeclareMathOperator*{\argmax}{argmax}
- \DeclareMathOperator*{\argmin}{argmin}
- \DeclareMathOperator{\sgn}{sgn}
---

@@

```{=latex}
\begin{equation}
\phi_{\text{BC}}^{\lambda, x_0, s} (x_i) = 
\begin{cases}
\left( \left(\frac{x_i - x_0}{s} \right)^\lambda - 1 \right) / \lambda & \text{if } \lambda \neq 0\\
\log\left[\frac{x_i - x_0}{s}\right] & \text{if } \lambda = 0
\end{cases}
\end{equation}
```
Where $x_i - x_0 > 0$.
The location and scale-invariant Yeo-Johnson transformation is:

```{=latex}
\begin{equation}
\phi_{\text{YJ}}^{\lambda, x_0, s} (x_i) = 
\begin{cases}
\left( \left( 1 + \frac{x_i - x_0}{s}\right)^\lambda - 1\right) / \lambda & \text{if } \lambda \neq 0 \text{ and } x_i - x_0 \geq 0\\
\log\left[1 + \frac{x_i - x_0}{s}\right] & \text{if } \lambda = 0 \text{ and } x_i - x_0 \geq 0\\
-\left( \left( 1 - \frac{x_i - x_0}{s}\right)^{2 - \lambda} - 1 \right) / \left(2 - \lambda \right) & \text{if } \lambda \neq 2 \text{ and } x_i - x_0 < 0\\
-\log\left[1 - \frac{x_i - x_0}{s}\right] & \text{if } \lambda = 2 \text{ and } x_i - x_0 < 0
\end{cases}
\end{equation}
```

@@

Since in our situation $\{x_1, \ldots, x_n\}$ in $f^{\lambda, x_0, s}(x_1, \ldots, x_n)$ are considered fixed (i.e., known), $f^{\lambda, x_0, s}(x_1, \ldots, x_n)$ may be considered a likelihood function.
The log-likelihood function $\mathcal{l}^{\lambda, x_0, s}$ is then:
```{=latex}
\begin{equation}
\begin{split}
\mathcal{l}^{\lambda, x_0, s} & = \log f^{\lambda, x_0, s}(x_1, \ldots, x_n) \\
 & = \log \left[ f^{\lambda, x_0, s} (\phi^{\lambda, x_0, s}(x_1), \ldots, \phi^{\lambda, x_0, s}(x_n)) \right] + \log \left|\mathbf{J}\right| \\
 & = \log \left[ f^{\lambda, x_0, s} (\phi^{\lambda, x_0, s}(x_1), \ldots, \phi^{\lambda, x_0, s}(x_n)) \right] + \log \prod_{i=1}^n \frac{\partial}{\partial x_i} \phi^{\lambda, x_0, s}(x_i) \\
 & = -\frac{n}{2} \log \left[2 \pi \sigma^2 \right] -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi^{\lambda, x_0, s}(x_i) - \mu \right)^2 + \sum_{i=1}^n \log \left[ \frac{\partial}{\partial x_i} \phi^{\lambda, x_0, s}(x_i)\right]
\end{split}
\end{equation}
```
With $\mu$ the average of $\phi^{\lambda, x_0, s}(\mathbf{X})$ and $\sigma^2$ its variance.
The first two terms derive directly from the log-likelihood function of a normal distribution, and are not specific to the type of power transformation used.
However, the final term differs between Box-Cox and Yeo-Johnson transformations.

@@

Thus the final term in $\mathcal{l}_{\text{BC}}^{\lambda, x_0, s}$ is:
```{=latex}
\begin{equation}
\begin{split}
\sum_{i=1}^n \log \frac{\partial}{\partial x_i} \phi_{\text{BC}}^{\lambda, x_0, s}(x_i) & = \sum_{i=1}^n \log \left[ s^{-\lambda} (x_i - x_0)^{\lambda - 1} \right] \\
& = \sum_{i=1}^n \log \left[s^{-\lambda} \right] + \log \left[ (x_i - x_0)^{\lambda - 1} \right]\\
& = -n \lambda \log s + \left( \lambda - 1 \right) \sum_{i=1}^n \log \left[ x_i - x_0 \right]
\end{split}
\end{equation}
```

This leads to the following log-likelihood for the Box-Cox transformation:
```{=latex}
\begin{equation}
\begin{split}
\mathcal{l}_{\text{BC}}^{\lambda, x_0, s} = & -\frac{n}{2} \log{2 \pi \sigma^2} -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi^{\lambda, x_0, s}(x_i) - \mu \right)^2 \\
& -n \lambda \log s + \left( \lambda - 1 \right) \sum_{i=1}^n \log \left[ x_i - x_0 \right]
\end{split}
\end{equation}
```

@@

Thus the final term in $\mathcal{l}_{\text{YJ}}^{\lambda, x_0, s}$ is:
```{=latex}
\begin{equation}
\begin{split}
\sum_{i=1}^n \log \frac{\partial}{\partial x_i} \phi_{\text{YJ}}^{\lambda, x_0, s}(x_i) & = - n \log s + (\lambda - 1) \sum_{i=1}^n \sgn(x_i - x_0) \log \left[1 + \frac{|x_i - x_0|}{s} \right]
\end{split}
\end{equation}
```

This lead to the following log-likelihood for the Yeo-Johnson transformation:
```{=latex}
\begin{equation}
\begin{split}
\mathcal{l}_{\text{BC}}^{\lambda, x_0, s} = & -\frac{n}{2} \log{2 \pi \sigma^2} -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi^{\lambda, x_0, s}(x_i) - \mu \right)^2 \\
& - n \log s + (\lambda - 1) \sum_{i=1}^n \sgn(x_i - x_0) \log \left[1 + \frac{|x_i - x_0|}{s} \right]
\end{split}
\end{equation}
```
